---
title: "Importing peaks"
shorttitle: "Import narrowPeak and broadPeak files as GRanges"
author:
  - name: "Charles Joly Beauparlant"
    affiliation: "Centre de Recherche du CHU de Québec - Université Laval, Québec, Québec, Canada, Département de Médecine Moléculaire, Faculté de médecine, Québec, Canada"
package: peakimport
abstract: >
  The goal of this package is to provide a function to easily import `narrowPeak` and `broadPeak` files.  The rationale is that there is currently a [bug](https://github.com/GuangchuangYu/ChIPseeker/pull/93) in the `r Biocpkg("ChIPseeker")` package. When `narrowPeak` files are used, the 1st line is used as an header. The first entry of the `narrowPeak` file is thus removed from the analysis. The same problem is found with `broadPeak` files. Until the bug is solved, it will be easier to load the `narrowPeak` and `broadPeak` files using this package. The `import_peaks` function can also be used in other context when `narrowPeak` and `broadPeak` files need to be imported.
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Annotating narrowPeak and broadPeak files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Vignette setup

For this vignette, we will import `narrowPeak` and `broadPeak` files. We will
then show how to annotate the results with the `r Biocpkg("ChIPseeker")`
package. To complete this demo, you will need the following packages:

```{r setup, message=FALSE}
library(peakimport)
require(ChIPseeker)
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
require(org.Hs.eg.db)
require(purrr)
```

# Importing files

In this section we will show how to import both `narrowPeak` and `broadPeak`
files. We will first show how to import each type separately, and then we will
show how we can use `lapply` or `map` to import multiple files with a single
command.

## `narrowPeak` files

The `narrowPeak` file format is described on the
[UCSC webpage]("http://genome.ucsc.edu/FAQ/FAQformat.html#format12").

It is not a valid format and cannot be directly imported using the `import`
function from `rtracklayer`.

```{r import_narrowPeak}
narrowPeak_file <- get_demo_file("narrowPeak")
gr_narrowPeak <- import_peaks(narrowPeak_file)
gr_narrowPeak
```

The results is a `GRanges` object.

## `broadPeak` files

The `broadPeak` file format is described on the
[UCSC webpage]("http://genome.ucsc.edu/FAQ/FAQformat.html#format13").

Similar to the `narrowPeak` format, it is not a valid `bed` file and will need
to be imported accordingly.

```{r import_broadPeak}
broadPeak_file <- get_demo_file("broadPeak")
gr_broadPeak <- import_peaks(broadPeak_file)
gr_broadPeak
```

The results is also a `GRanges` object.

## Importing multiple files

We can use functions such as `lapply` or `map` (from the `r CRANpkg("purrr")`
package) to import multiple files in a single command:

```{r import_multiple_files}
multiple_files <- c(narrowPeak_file, broadPeak_file)
names(multiple_files) <- c("narrowPeak", "broadPeak")
grl <- lapply(multiple_files, import_peaks)
# or
grl <- map(multiple_files, import_peaks)
grl
```

Both `lapply` and `map` will return the same result: a `list` of `GRanges`.

The `import_peaks` function will determine if the file is a `bed`, `narrowPeak`
or `broadPeak` file based on the file extension and will select the correct
import strategy accordingly. It is important for the file extensions to be
exactly `.narrowPeak` and `.broadPeak`.

# Annotating results

We will use the `r Biocpkg("ChIPseeker")` package to annotate the results from
`import_peaks`:

```{r annotate_peaks}
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

anno_narrowPeak <- annotatePeak(gr_narrowPeak,
				TxDb = txdb,
				annoDb = "org.Hs.eg.db")
anno_narrowPeak
anno_broadPeak <- annotatePeak(gr_broadPeak,
				TxDb = txdb,
				annoDb = "org.Hs.eg.db")
anno_broadPeak
```

The result can be converted in the `GRanges` format with the `as.GRanges`
function:

```{r convert_to_GRanges}
gr_anno_narrowPeak <- as.GRanges(anno_narrowPeak)
gr_anno_narrowPeak
gr_anno_broadPeak <- as.GRanges(anno_broadPeak)
gr_anno_broadPeak
```

Finally, we can annotate multiple peaks in a single command with `lapply` or
`map`:

```{r annotate_multiple_files}
anno_grl <- map(grl, annotatePeak, TxDb = txdb, annoDb = "org.Hs.eg.db")
gr_anno_grl <- map(anno_grl, as.GRanges)
gr_anno_grl
```

# `sessionInfo()`

```{r sessionInfo}
sessionInfo()
```
